---
title: "Data Analysis from GPS Collar Prototype"
output: github_document
---
Analysis of properties of the GPS data from the prototype GPS collar
12 March 2018

With the prototype GPS and the firmware dated 3/11/2018, I conducted a series of tests to empirically measure 1) power consumption over a longer time period (24 hr) to get a handle on battery capacity, 2) time to first average fix and time to best fix, and 3) quality of the GPS readings in terms of average HDOP and spread around the average GPS reading from a stationary prototype.

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(rgdal)
library(dplyr)
library(geosphere)
```


```{r data, echo=FALSE}
# Load Data from COTS prototype
path <- "C:\\Users\\Jason Karl\\Documents\\GitHub\\COTS_GPS_Collars\\COTS_GPS_Collars\\GPS_Evaluation"
rtk.loc <- c(-116.991833249,46.7172203903)


file <-"point_averaging_5-sec-niterval_GPSLOG.CSV" # "1sec_test2_1hr_duration.csv" #
colNames <- c("HDOP","Lat","Lon","NumSat","UTCDate","UTCTime","FirstFix","LastFix","BestFix")
gps.data <- read.csv(paste(path,file,sep="\\"),header=T,col.names=colNames,stringsAsFactors=F)

# Filter to remove blanks
gps.data <- gps.data[gps.data$Lon<0,]
gps.data <- gps.data[gps.data$NumSat>0,]

## calculate mean lat/lon
meanLat <- mean(gps.data$Lat)
meanLon <- mean(gps.data$Lon)

## Calc difference from mean
gps.data$diffdist <- mapply(function(x,y){distGeo(c(meanLon,meanLat),c(x,y))},x=gps.data$Lon,y=gps.data$Lat)
gps.data$diffLat <- mapply(function(x,y){distGeo(c(x,meanLat),c(x,y))},x=gps.data$Lon,y=gps.data$Lat)
gps.data$diffLon <- mapply(function(x,y){distGeo(c(meanLon,y),c(x,y))},x=gps.data$Lon,y=gps.data$Lat)


## Load the iGotU data
iGotU.file <- "iGotU_unit5_test.csv"
iGotU.data <- read.csv(paste(path,iGotU.file,sep="\\"),header=T,stringsAsFactors=F,fileEncoding = "UTF-16LE")

## remove the blanks from the iGotU data and filter to one day
iGotU.data <- iGotU.data %>% filter(Latitude > 0 & Date == ' 2018/03/25')

## find mean of iGotU GPS locations
iGotU.meanLat <- mean(iGotU.data$Lat)
iGotU.meanLon <- mean(iGotU.data$Lon)

## Calc difference from mean iGotU loc
iGotU.data$diffdist <- mapply(function(x,y){distGeo(c(iGotU.meanLon,iGotU.meanLat),c(x,y))},
                              x=iGotU.data$Lon,y=iGotU.data$Lat)
iGotU.data$diffLat <- mapply(function(x,y){distGeo(c(x,iGotU.meanLat),c(x,y))},
                              x=iGotU.data$Lon,y=iGotU.data$Lat)
iGotU.data$diffLon <- mapply(function(x,y){distGeo(c(iGotU.meanLon,y),c(x,y))},
                              x=iGotU.data$Lon,y=iGotU.data$Lat)

```
## Summary Statistics for GPS Test Data


```{r summaries, echo=FALSE}
## Calc diference from RTK location to mean of collar
gps.accuracy <- distGeo(rtk.loc,c(meanLon,meanLat))
iGotU.accuracy <- distGeo(rtk.loc,c(iGotU.meanLon,iGotU.meanLat))

gps.95CEP <- quantile(gps.data$diffdist,0.95)
iGotU.95CEP <- quantile(iGotU.data$diffdist,0.95)
```



```{r plots}

g <- ggplot(data=gps.data, aes(x=diffLon,y=diffLat))+geom_point(color="grey")+
  stat_ellipse(level=0.95)+ #geom_hline(yintercept = meanLat) + geom_vline(xintercept = meanLon)+ #geom_point(aes(x=rtk.loc[1],y=rtk.loc[2]))+
  labs(title="COTS GPS positional error", x="Difference from Avg Longitude (m)", y="Difference from Avg Latitude (m)")
g

g.iGotU <- ggplot(data=iGotU.data, aes(x=diffLon,y=diffLat))+geom_point(color="grey")+
  stat_ellipse(level=0.95)+geom_hline(yintercept = 0) + geom_vline(xintercept = 0)+
  xlim(c(-50,75))+ylim(c(-50,50))+
  labs(title="iGotU GPS positional error", x="Difference from Avg Longitude (m)", y="Difference from Avg Latitude (m)")
g.iGotU
```

### Trend of positional accuracy by HDOP?
Yeah, but pretty slight.
```{r}
ggplot(data=plot.df[plot.df$HDOP<500,], aes(x=HDOP, y=diffdist))+geom_point()+geom_smooth(method="lm")
```


### Is there a trend of positional error by time of day?
Nope.

```{r}
ggplot(data=plot.df, aes(x=GPSTime, y=diffdist))+geom_point()
```
```{r, include=FALSE}
print(mean(plot.df$FixTime))
print(quantile(plot.df$FixTime, 0.9))

print(mean(plot.df$BestTime))
print(quantile(plot.df$BestTime, 0.9))

print(mean(plot.df$diffdist))
print(quantile(plot.df$diffdist, 0.9))
```

## Takehome messages
For 10-minute fix intervals with the GPS unit inside the house (I'm kind of surprised it even got a fix):
- Average time to first fix was 11 seconds with 90% of first fixes within 27.3 seconds.
- Average time to best fix was 13 seconds with 90% of first fixes within 31.3 seconds.
- Average positional error was 14.4m with 90% of GPS readings within 29.4m of the average of all GPS readings.



